Waking a sleeping rabbit:\\ On Natural-Language Sentence Generation with FF

### 2009-12-18 trunk/Papers/crisp-ff-10/experiments.tex
experiments v2, added pictures

.. describe before/after picture, claim that now the performance is
``for the moment ok for this application'', but not for foreseeable
future (bigger grammers etc)

>>>>

To evaluate the effect of our changes on the efficiency of the planner
on practical inputs, we ran a series of experiments using the XTAG
Grammar \cite{xtag01:_tr}, a large-scale tree-adjoining grammar of
English, from which we selected all lexicon entries for seven words we
used in the experiments. We set up inputs to generate sentences of the
form ``$S_1$ and $S_2$ and \ldots and $S_n$'', i.e.\ a conjunction of
$n$ sentences. Each sentence $S_i$ is of the form ``X admires Y''. X
and Y are unique referring expressions to individuals, such as ``the
rich businessman''. For each experiment, we have a parameter $d$ that
specifies how many adjectives are needed to distinguish the individual
uniquely -- in particular, $d=0$ means we can say ``the businessman'',
and $d=2$ means we must say ``the rich sad businessman''. We translate
these generation problems into planning problems with the new encoding
as shown in Fig.~\ref{fig:white-rabbit-as-planning}.

\begin{figure}[t]
  \centering
  \includegraphics[width=\columnwidth]{pics/xtag-k2-dist0}
  \caption{Runtimes for $d=0$.}
  \label{fig:runtimes-k2-dist0}
\end{figure}

\begin{figure}[t]
  \centering
  \includegraphics[width=\columnwidth]{pics/xtag-k2-dist2}
  \caption{Runtimes for $d=2$.}
  \label{fig:runtimes-k2-dist2}
\end{figure}

Consider first the results for $d=0$, which are shown in
Fig.~\ref{fig:runtimes-k2-dist0}; the lines end where the planner
exceeded its five-minute timeout.  The top line in the figure
represents the off-the-shelf version of FF. The other two lines
represent FF with the improved preprocessor, using the standard
enforced hillclimbing strategy and the best-first search with helpful
actions (``New FF'') respectively.  By improving the preprocessor, we
have improved its runtime on small instances by three orders of
magnitude; this factor only increases for larger inputs.  After the
preprocessor has been improved, both search strategies can solve these
problem instances in practically useful time -- for instance, the
24-word sentence at $n=4$ is generated in under two seconds.

To get a clearer view of the relative efficiency of the two search
strategies, we also ran the experiment with $d=2$; the results are
shown in Fig.~\ref{fig:runtimes-k2-dist2}.  Here the old FF
preprocessor crashed with a segmentation fault even at
$n=1$. Best-first search with helpful actions is clearly the
best-performing search strategy here.  It takes about 14 seconds for
$n=3$, but this sentence is already quite long at 29 words, and EHC
takes 3.5 minutes on the same input.  Best-first search without
helpful actions timed out at $n=3$, which illustrates the advantage we
gain from using the helpful actions heuristic.

While this brings the overall planner runtimes into a range where FF
can now be useful in practical NLG applications, it is important to
note that it still has serious limitations.  The best-first search
timed out for any $n>4$ in either experiment.  While many NLG
applications can get by with shorter sentences, improving the search
on our domain is still an open question for the future.  Even within
our domain, the relative quality of EHC and BFS+H depended on the
exact generation problem instance: While BFS+H outperformed EHC at
$d=2$, EHC was actually slightly faster at $d=0$ because the absence
of distractors meant there was not much room for the characteristic
underestimation of the relaxed plan evaluation discussed above.  This
means that sentence generation remains as a varied and challenging
domain for planning.
